Measurement Bias from Address Aliasing

### 2014-05-13 paper.tex
Improved handling of raw data for convolution example
More sane way of importing and drawing table/graph

% Detailed comparison between median and worst cases.
\begin{table}[h]
  \caption{Events with significant correlation to cycle count.}
  \label{tab:loopcorrelation}
  \centering

  \renewcommand{\tabcolsep}{1pt} % Less padding between cells
  \pgfkeys{/pgf/number format/.cd,int trunc} % Output whole numbers

  \pgfplotstabletypeset[
    font=\scriptsize,
    col sep=comma,
    columns={Performance counter, Median, [index]3, [index]4},
    column type=r,
    columns/Performance counter/.style={
      string type, 
      column type=l,
      postproc cell content/.append code={
        \pgfkeysalso{@cell content=\perfctr{##1}}
      }
    },
    every head row/.style={
      output empty row,
      before row={\toprule
        Performance counter & Median & Spike 1 & Spike 2 \\
        },
      after row=\midrule,
    },
    every last row/.style={after row=\bottomrule}
  ]{bin/micro-kernel-comparison.csv}
\end{table}

@@

\section{Bias from Heap Allocation}
% Intro
% - Bias from aliasing caused by accesses to any part of memory
% - Dynamic memory usually from dynamically linked libs = bias
% - More interesting: default behavior provokes aliasing 
% Heap allocation alias by default
% - Normal heap with brk/sbrk syscalls
%   - Manage heap and increase on demand
% - Different depending on size (and other parameters)
%   - Use mmap for performance, usually large allocations
%   - Posix function, anonymous memory mapping
%   - Page aligned
% - Normal addr. vs mmap addr.
% - Makes no difference ++ everything is conceptually the same heap, but alias by default!
% Aligned sequential access
% - Intel warning, copy routine. 
% - Trouble with aligned read/write buffers, creating constant conflicts in sliding window.
% - More contrived example; convolution
%   - Simple code disregarding endpoints
%   - Manipulate start addresses by offseting pointers
%   - Compiled with gcc O3, shows bad performance (graph) + discussion
% - Table with lots of data points, showing actual numbers for each offset
% Architectural optimization
% - Modified example with restrict and -march=i7
% Bahavior of other Allocators
% - Intro each allocator
% - Present table
% Conclude
% - not good...
% - Not tested other compilers, not the purpose of this paper. 
% - How did we find this example? Intelligent design, literally first guess

@@

\subsection{Aligned Sequential Access}
Consider the function shown in Listing~\ref{lst:conv}.
It computes the convolution between an input array and a fixed five-element kernel, writing the result to another array.
For simplicity, endpoints are not handled.

% Plot of cycles vs alias for estimated cost of O3 convolution
\pgfplotstableread{bin/default-o3.estimate.dat}{\convolutiontable}
\begin{figure}[t]
  \caption{Cycle and alias counts for different offsets between input and output arrays in Listing~\ref{lst:conv}. Offset 0 means equal 12 bit address suffix, which is default behavior for mmap allocations. Array size $n=2^{20}$, compiled with -O3. }
  \label{fig:heapalias}
  \begin{tikzpicture}
    \begin{axis}[
        font = \small,
        xlabel=Relative offset in \lstinline!sizeof(float)! bytes,
        cycle list name=black white
      ]
      \addplot table[x expr = \thisrowno{0}, y = cycles:u] \convolutiontable ;
      \addplot table[x expr = \thisrowno{0}, y = r0107:u ] \convolutiontable ;
      \addlegendentry{cycles:u} ;
      \addlegendentry{r0107:u} ;
    \end{axis}
  \end{tikzpicture}
\end{figure}

\begin{lstlisting}[float=t, language=C, caption={Naive implementation of convolution. Highly sensitive to aliasing between input and output arrays.}, label={lst:conv}, frame=lines]
static float k[5]={0.1,0.25,0.3,0.25,0.1};
void conv(int n, float *in, float *out) {
    int i, j;
    for (i = 2; i < n-2; ++i) {
      out[i] = 0;
      for (j = 0; j < 5; ++j)
          out[i] += in[i-2+j] * k[j];
    }
}
\end{lstlisting}

@@

% Table with detailed performance data, estimated for O3
\pgfplotstabletranspose[
  col sep=comma,
  columns={n, cycles:u,r0107:u,r02a3:u,r01a2:u,r05a3:u,r04a2:u,r20a1:u,r0860:u,r0160:u,r2049:u,r6049:u},
  colnames from=n % Automapped to new 'colnames' column
  ]\loadedtable{bin/default-o3.estimate.dat}

\begin{sidewaystable*}[p]
  \caption{Most interesting performance counters for O3 convolution code, estimated cost accounting for constant overhead of heap allocation in main.}
  \label{tab:convstats}
  \pgfkeys{/pgf/number format/.cd,int trunc} % Output whole numbers
  \pgfplotstabletypeset[
    font=\normalsize,
    columns={[index]0, [index]1, [index]2, [index]3, [index]4, [index]5, [index]11, [index]21, [index]31}, % Select interesting offsets
    columns/colnames/.style={string type,column name={}, column type={l}},
    columns/colnames/.append style={string replace={cycles:u}{Cycles}}, % Pretty print counter mnemonics
    columns/colnames/.append style={string replace={r0107:u}{Alias}},
    columns/colnames/.append style={string replace={r02a3:u}{cycle_activity.cycles_ldm_pending}},
    columns/colnames/.append style={string replace={r01a2:u}{resource_stalls.any}},
    columns/colnames/.append style={string replace={r05a3:u}{cycle_activity.stalls_l2_pending}},
    columns/colnames/.append style={string replace={r04a2:u}{resource_stalls.rs}},
    columns/colnames/.append style={string replace={r20a1:u}{uops_executed_port.port_5}},
    columns/colnames/.append style={string replace={r0860:u}{offcore_requests_outstanding.all_data_rd}},
    columns/colnames/.append style={string replace={r0160:u}{offcore_requests_outstanding.demand_data_rd}},
    columns/colnames/.append style={string replace={r2049:u}{dtlb_store_misses.stlb_hit_4k}},
    columns/colnames/.append style={string replace={r6049:u}{dtlb_store_misses.stlb_hit}},
    every head row/.style={before row=\toprule, after row=\midrule},
    every last row/.style={after row=\bottomrule}
    ]{\loadedtable}
\end{sidewaystable*}

>>>>

% Detailed comparison between median and worst cases.
\begin{table}[h]
  \caption{Events with significant correlation to cycle count.}
  \label{tab:loopcorrelation}
  \centering
  \renewcommand{\tabcolsep}{1pt} % Less padding between cells
  \pgfplotstabletypeset[
    font=\scriptsize,
    int detect, % Output whole numbers for counter values
    col sep=comma,
    columns={Performance counter, Median, [index]3, [index]4},
    column type=r,
    columns/Performance counter/.style={
      string type, 
      column type=l,
      postproc cell content/.append code={
        \pgfkeysalso{@cell content=\perfctr{##1}}
      }
    },
    every head row/.style={
      output empty row,
      before row={\toprule
        Performance counter & Median & Spike 1 & Spike 2 \\
        },
      after row=\midrule,
    },
    every last row/.style={after row=\bottomrule}
  ]{bin/micro-kernel-comparison.csv}
\end{table}

@@

\section{Bias from Heap Allocation}
% Intro
% - Bias from aliasing caused by accesses to any part of memory
% - Dynamic memory usually from dynamically linked libs = bias
% - More interesting: default behavior provokes aliasing 
% - Intel warning, copy routine. 
% - Trouble with aligned read/write buffers, creating constant conflicts in sliding window.
% Heap allocation alias by default
% - Normal heap with brk/sbrk syscalls
%   - Manage heap and increase on demand
% - Different depending on size (and other parameters)
%   - Use mmap for performance, usually large allocations
%   - Posix function, anonymous memory mapping
%   - Page aligned
% - Normal addr. vs mmap addr.
% - Makes no difference ++ everything is conceptually the same heap, but alias by default!
% Convolution kernel Analysis
% - More contrived example; convolution
%   - Simple code disregarding endpoints
%   - Manipulate start addresses by offseting pointers
%   - Compiled with gcc O3, shows bad performance (graph) + discussion
% - Table with lots of data points, showing actual numbers for each offset
% Architectural optimization
% - Modified example with restrict and -march=native
% Bahavior of other Allocators
% - Intro each allocator
% - Present table
% Conclude
% - not good...
% - Not tested other compilers, not the purpose of this paper. 
% - How did we find this example? Intelligent design, literally first guess

@@

\subsection{Analysis of Convolution kernel}
Consider the function shown in Listing~\ref{lst:conv}.
It computes the convolution between an input array and a fixed five-element kernel, writing the result to another array.
For simplicity, endpoints are not handled.

% Plot of cycles vs alias for estimated cost of O3 convolution
\pgfplotstableread{bin/conv-default-o3.estimate.dat}{\convolutiontable}
\begin{figure}[t]
  \caption{Cycle and alias counts for different offsets between input and output arrays in Listing~\ref{lst:conv}. Offset 0 means equal 12 bit address suffix, which is default behavior for mmap allocations. Array size $n=2^{20}$, compiled with -O3. }
  \label{fig:heapalias}
  \begin{tikzpicture}
    \begin{axis}[
        font = \small,
        xlabel=Relative offset in \lstinline!sizeof(float)! bytes,
        cycle list name=black white
      ]
      \addplot table[x expr = \thisrowno{0}, y = cycles:u] \convolutiontable ;
      \addplot table[x expr = \thisrowno{0}, y = r0107:u ] \convolutiontable ;
      \addlegendentry{cycles:u} ;
      \addlegendentry{r0107:u} ;
    \end{axis}
  \end{tikzpicture}
\end{figure}

\lstinputlisting[
  float=t, 
  language=C, 
  caption={Naive implementation of convolution. Highly sensitive to aliasing between input and output arrays.},
  label={lst:conv},
  frame=lines
]{bin/convolution-kernel.c}

@@

% Table with detailed performance data, estimated for O3 (sidewaystable)
\begin{table*}[t]
  \caption{Most interesting performance counters for O3 convolution code, estimated cost accounting for constant overhead of heap allocation in main.}
  \label{tab:convstats}

  \renewcommand{\tabcolsep}{3pt} % Less padding between cells
  \pgfplotstabletypeset[
    font=\footnotesize,
    int detect, % Output whole numbers for counter values
    col sep=comma,
    columns={Performance counter, Correlation, 0, 2, 4, 8, 16, 24},
    column type=r,
    columns/Performance counter/.style={
      string type, 
      column type=l,
      postproc cell content/.append code={
        \pgfkeysalso{@cell content=\perfctr{##1}}
      }
    },
    columns/Correlation/.style={
      fixed,
      fixed zerofill,
      precision=2
    },
    every head row/.style={
      output empty row,
      before row={\toprule
        Performance counter & $r$ & 0 & 2 & 4 & 8 & 16 & 24 \\
        },
      after row=\midrule,
    },
    every last row/.style={after row=\bottomrule}
  ]{bin/conv-default-o3.estimate.csv}
\end{table*}
